### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1212f7fb3388b045df17ec4f4ef36d65913d0183 🚀

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{38,39} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with more
than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
broad variety of habitat use, behaviour, and feeding strategies, putting
them at key positions in the delivery and provisioning of several
ecosystem services, tied to important ecosystem-derived benefits to
humans.\textsuperscript{42} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{43,44} and
vectors of pathogens that put a risk on human
health;\textsuperscript{45,46} other species are essential links in many
seed-dispersal networks.\textsuperscript{47} Areas where bats, viruses,
and humans co-occur are not necessarily hotspots of risk for human
heath; as such, developing more precise ways to map zoonotic hazards can
help bats and humans coexist safely and support the conservation of
these important and unique animals.

@@

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,49} Bats---and the spillover of their
viruses---are also sensitive responds to anthropogenic factors others
than shifting climates, including deforestation and other kinds of
habitat loss, increased stress, and greater contact with potential
bridge hosts like domesticated species.\textsuperscript{26,50,51} This
represents a challenge for both conservation strategies and pandemic
prevention,\textsuperscript{52} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both (add cites for DOIs:
10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).

>>>>

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{38,39} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with more
than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
broad variety of habitat use, behaviour, and feeding strategies, putting
them at key positions in the delivery and provisioning of several
ecosystem services, tied to important ecosystem-derived benefits to
humans.\textsuperscript{42} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{43,44} and
vectors of pathogens that put a risk on human
health;\textsuperscript{45,46} some other species are essential links in
many seed-dispersal networks.\textsuperscript{47} However, many of these
species face a high risk of extinction, particularly given persecution
and killings that sometimes follows from messaging about their role in
disease emergence. Areas where bats, viruses, and humans co-occur are
not always hotspots of risk for human heath; as such, developing more
precise ways to map zoonotic hazards can help bats and humans coexist
safely, and support the conservation of these important and unique
animals.

@@

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{28,49} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{26,50,51} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{52} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both (add cites for DOIs:
10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).
